CUBE: A Memory-optimized Cosmological $N$-body Algorithm

### 2017-10-19 ms_caf/ms_caf.tex
draft v1.0

\shortauthors{Yu et al.} \shorttitle{CUBE}
\begin{document}
\title{CUBE: A Memory-lite Cosmological $N$-body Algorithm}
\author{
Hao-Ran Yu$^{1,2,3}$,
Ue-Li Pen$^{2,4,5,6}$,
Xin Wang$^{2}$
}

@@

When $N$ is large, the brute force pairwise particle-particle (PP) force brings unaffordable $o(N^2)$ computations. The workload can be alleviated by many algorithms, for example, the ``tree'' algorithms like {\tt GADGET} \citep{2001NewA....6...79S,2005MNRAS.364.1105S} and its simulation ``Millennium'' \citep{2005Natur.435..629S,2012MNRAS.426.2046A}, {\tt TPM} \citep{1995ApJS...98..355X} and {\tt GOTPM} \citep{2004NewA....9..111D}, adaptive grid algorithm including {\tt HYDRA} \citep{1995ApJ...452..797C} and {\tt RAMSES} \citep{2010ascl.soft11007T}, as well as mesh-refined codes \citep{1991ApJ...368L..23C} and moving adaptive particle-mesh codes \citep{1995ApJS..100..269P}. Among these, the standard particle-mesh (PM) algorithm \citep{1988csup.book.....H} is most memory and computational efficient if we focus on large cosmological scales. The load balancing problem is minimized as the matter distribution is rather homogeneous, and the speed benefits from the fast Fourier transform (FFT) libraries, such as {\tt FFTW3} \citep{Frigo05thedesign}. Based on PM, the parallel code {\tt PMFAST} \citep{2005NewA...10..393M} uses a two-level PM algorithm to reduce the global communication and memory. After subsequent developments, {\tt CUBEP3M} \citep{2013MNRAS.436..540H} uses cubic spatial decomposition, and adds in-cell or adjacent-cell PP force and many other features.

With the fast development of parallel 
supercomputers, we are able to simulate a system of more than a trillion ($10^{12}$) 
$N$-body particles. To date the largest $N$-body simulation in application is the 
``TianNu'' simulation \citep{2017NatAs...1E.143Y,2017RAA....17...85E} run on the 
TianHe-2 supercomputer by cosmological simulation code 
{\tt CUBEP3M} \citep{2013MNRAS.436..540H}. It uses nearly $3\times 10^{12}$ particles 
to simulate the cold dark matter (CDM) and cosmic neutrino evolution through the 
cosmic age.

$N$-body simulations use considerable amount of memory, because the phase space 
coordinates $(x,y,z,v_x,v_y,v_z)$ of each $N$-body particle must be stored as, at 
least, six single precision floating point numbers (24 bytes). Contrarily, their 
computing workload can be alleviated by many algorithms, like Particle-Mesh [cite] and 
Tree [cite], scale as $o(N\log N)$ or even $o(N)$. On the other hand, modern 
supercomputer systems use multi cores, many integrated cores (MIC) and even densely 
parallelized GPUs, bringing orders of magnitude higher computing power, whereas these 
architectures usually have limited memory allocation. Thus, the computation-light but 
memory-heavy applications, compared to matrix multiplication and decomposition 
calculations, are less suitable for fully usage of the computing power of modern 
supercomputers. For example, although native and offload modes of {\tt CUBEP3M} are 
able to run on the Intel Xeon-PHI MIC architectures, with the require ment of enough 
memory, TianNu simulation were done on TianHe-2 with only its CPUs -- 73\% of the 
total memory but only 13\% of the total computing power.

We present a new $N$-body simulation code {\tt CUBE}, using as low as 6 byte
per particle (here after we use ``bpp'' referring to ``byte per particle'').
We show that it gives accurate results in cosmological LSS simulations.
The method is presented in \S\ref{s.method}, and a comparison between this method
and traditional method is shown in \S\ref{s.results}. Discussions and 
conclusions are in \S\ref{s.discussion}.

\section{Method}\label{s.method}
The most memory consuming part of a $N$-body simulation is usually the phase space 
coordinates of $N$-body particles -- 24 bpp (4 single precision floating numbers) must 
be used to store each particles' 3D position and velocity vectors. {\tt CUBEP3M}, an 
example of a memory-efficient parallel $N$-body code, can use as low as 40 bpp in 
sacrificing computing speed \citep{2013MNRAS.436..540H}. This includes the phase 
coordinates (24 bpp) for particles in physical domain and buffered region, a linked 
list (4 bpp), and a global coarse mesh and local fine mesh. 4-byte real numbers are 
not necessarily adequate in representing the {\it global} coordinates in simulations.
If the box size is many orders of magnitude larger than interactive distance
between particles, especially in the field of re-simulation of dense subregions,
double precision (8-byte) coordinates are needed to avoid truncation errors.
Another solution is to record {\it relative} coordinates for both position and velocity.
{\tt CUBE} replaces the coordinates and linked list 24+4=28 bpp memory usage
with an integer based storage, reduces the basic memory usage from 28 bpp down to
6 bpp, described as following \ref{ss.position} and \ref{ss.velocity}. The algorithm is described in \ref{ss.overview}.

\subsection{Particle position storage}\label{ss.position}
We construct a uniform mesh throughout the space and each particle belongs 
to its parent cell of the mesh. Instead of storing global coordinates of 
each particle, we store its offset relative to its parent cell which 
contains the particle. We divide the cell, in each dimension $d$, evenly into 
$2^8=256$ bins, and use a 1-byte (8 bits) integer $\chi_d \in \{-128,-127,...,127\}$ to 
indicate which bin it locates in this dimension. The global locations of 
particles are given by cell-ordered format in memory space, and a 
complimentary number count of particle number in this mesh (density field) 
will give complete information of particle distribution in the mesh. Then 
the global coordinate in $d$th dimension $x_d$ is given by $x_d=(n_c-1)+
(\chi_d+128+1/2)/256$, where $n_c=1,2,...,N_c$ 
is the index of the coarse grid. The mesh 
is chosen to be coarse enough such that the density field takes negligible 
memory. This coarse density field can be further compressed into 1-byte 
integer format, such that a 1-byte integer show the particle number in this 
coarse cell in range 0 to 255. In the densest cells (rarely happened) where 
there are $\ge 255$ particles, we can just write 255, and write the actual 
number as a 4-byte integer in another file.

In a simulation with volume $L^3$ and $N_c^3$ coarse cells, particle 
positions are stored with a precision of $L/(256N_c)$. The force calculation 
(e.g. softening length) should be configured much finer than this 
resolution, discussed in later sections. On the other hand, particle 
position can also be stored as 2-byte (16 bits) integers to increase the resolution. 
In this case, each coarse cell is divided into $2^{16}=65536$ bins and the 
position precision is $L/(65536N_c)$, precise enough compared to using
4-byte global coordinates, see later results. We denote this case ``x2'' and 
denote using 1-byte integers for positions ``x1''.

@@

\begin{figure}
\centering
  \includegraphics[width=1.0\linewidth]{vdisp.pdf}
 \caption{Variance of the velocity distribution given by Equation \ref{eq.vdisp}. The $x$-axis is the scale factor characterizing the cosmic evolution and the $y$-axis shows the $\sigma_v$ in unit of km/s. The 5 curves, from top to bottom, shows the $\sigma_v(a)$ with increasing smoothing scale $r$ from 0 to 20 Mpc$/h$.}
\label{fig.vdisp}
\end{figure}

@@

The reason why we use uneven bins is that, slower particles are more 
abundant compared to faster ones, and one should better resolve slower 
particles tracing at least linear evolution. On the other hand, there could 
be extreme scattering particles (in case of particle-particle force), and we 
can safely ignore or less resolve those nonphysical particles. One of the 
solution is that, if we know the probability distribution function (PDF) 
$f(\Delta v)$ we divide its cumulative distribution function (CDF) $F(\Delta 
v)\in(0,1)$ into $2^{8n_\nu}$ bins to determine the boundary of $\Delta v$ 
bins, and particles should evenly distribute in the corresponding uneven $
\Delta v$ bins. Practically we find that either $f(v_d)$ or $f(\Delta v)$ is 
close to Gaussian, so we can use Gaussian CDF, or any convenient analytic 
functions which are close to Gaussian, to convert velocity between real 
numbers and integers.

The essential parameter of the velocity distribution is its variance. On 
non-linear scale, the velocity distribution function is non-Gaussian. 
However, to the first order approximation, we simply assume it as Gaussian 
and characterized by the variance
\begin{equation}\label{eq.vdisp}
	\sigma_v(a,k) = \frac{1}{3}  (a H f D)^2 \int_0^kd^3q\frac{P_{L}(q)}{q^2},
\end{equation}
where $a(z)$ is the scale factor, $H(z)$ the Hubble parameter, $D$ is the 
linear growth factor, $f=d \ln D/d\ln a$, and $P_L(k)$ is the linear power 
spectrum of density contrast at redshift zero.
$\sigma_v(a,k)$ is a function of cosmic evolution $a$ and a smoothing scale 
$k$, or $r$ (see Figure \ref{fig.vdisp}). $\Delta v$
is the velocity dispersion relative to the coarse grid, so we approximate 
its variance as
\begin{equation}\label{eq.vdelta}
	\sigma^2_{\Delta}(a)=\sigma^2_v(a,r_c)-\sigma^2_v(a,r_p), 
\end{equation}
where $r_c$ is the scale of
coarse grid, and $r_p$ is the scale of average particle separation. In each
dimension of 3D velocity field, we use $\sigma^2_{\Delta}(a)/3$ according 
to the equipartition therorem. On different scales, we measure the 
statistics of $v_d$, $v_c$ and $\Delta v$ and find good agreement with the 
above model.

\begin{figure}
\centering
  \includegraphics[width=0.99\linewidth]{fig_tile.pdf}
 \caption{Spacial decomposition in {\tt CUBE} in a 2D analogy. In this example, there are 2 images per dimension (${\tt nn}=2$), and 2 tiles per image per dimension (${\tt nnt}=2$). The orange boxes show the overlapped physical+buffer=extended tile regions, inside of which one physical regions is indicated with green.}
\label{fig.tile}
\end{figure}

@@

@@

{\tt update\_vp} -- 
The particle-mesh (PM) or particle-particle particle-mesh (P$^3$M) algorithm is applied in this subroutine to update particles's velocities.
After {\tt update\_xp}, or drift $D$, we call {\tt buffer\_density} and {\tt buffer\_xp} in order that particle positions are in buffered state, based on which, we update velocities (kick $K$) of particles in physical region of each tile. {\tt CUBE} uses a 2-level particle mesh scheme \citep{2013MNRAS.436..540H}. Local fine forces have a force cutoff ${\tt nforce\_cutoff}\leq{\tt ncb}$, i.e., $F_{\rm fine}(r>{\tt ncb})=0$. So, if we apply a fine-grid particle-mesh on an extended tile with buffered-depth {\tt ncb} (see Figure \ref{fig.tile} and regard it periodic), the force on physical regions does not depend on the false periodic boundary assumption, and the resulting fine force $F_{\rm fine}$ and velocity update on physical region is correct.

@@

\begin{figure*}[t]
\centering
  \includegraphics[width=0.75\linewidth]{overplotparticles.pdf}
 \caption{Offset in the particle distribution induced by integer-1 based algorithm (x1v1). In S512's subregion of volume $15.625^2\times 3.90625\ ({\rm Mpc}/h)^3$, particles from {\tt CUBEP3M} (bigger grey dots in the background) and {\tt CUBE}-x1v1 (smaller red dots) are projected onto the plane of $(15.625\ {\rm Mpc}/h)^2$. The comparing rules show 1 ${\rm Mpc}/h$, fine grid and coarse grid respectively, and position resolution of x1v1 is 1/64 of a fine grid.}
\label{fig.particles}
\end{figure*}

\subsubsection{Compare with traditional algorithms}
To illustrate the improvement of memory usage, we give an example of TianNu simulation \citep{2017NatAs...1E.143Y} on the Tianhe-2 supercomputer, which used a traditional {\tt CUBEP3M} code. TianNu's particles number is limited by memory per computing node -- per computing node, an average of $576^3$ neutrino particles and $288^3$ CDM particles are used, and consumes about 40 GB of memory\footnote{Additional memory is used for OpenMP parallelization and for particle IDs to differentiate different particle spices.}, or about 186 bpp. A memory efficient case of {\tt CUBEP3M} by using large physical scales and at costs of speed, still uses about 40 bpp.

By using {\tt CUBE}, with parameters $n_\chi=n_\nu=1$, ${\tt nc}=384$, ${\tt nnt}=3$, ${\tt ncb}=6$, $\epsilon_{\rm image}=5\%$, $\epsilon_{\rm tile}=200\%$ and $\left\langle {\tt nplocal} \right\rangle=1536^3$, we use about 35.6 GB of memory, corresponding to 9.84 bpp. This can be done on most of the supercomputers, even modern laptops.

@@

\subsection{Displacement of particles}
In S512, we zoom into a small region of $15.625^2\times 3.90625\ ({\rm Mpc}/h)^3$ and compare the particle distribution between {\tt CUBEP3M} and {\tt CUBE}-x1v1 in Figure \ref{fig.particles}. For clarity, {\tt CUBEP3M} particles are marked with bigger, grey dots, whereas smaller red dots are {\tt CUBE}-x1v1 particles over-plotted onto them. One can see the difference between them. The 1 ${\rm Mpc}/h$, fine grid and coarse grid scales are shown in the figure. The position resolution of particles in {\tt CUBE}-x1v1 is 1/256 of a coarse grid, or 1/64 of a fine grid.
\begin{figure}[]
\centering
  \includegraphics[width=0.95\linewidth]{ddsp_pcdf_256_box80.pdf}
 \caption{Statistics of the error in particle displacement $\Delta{\bs\Psi}$ induced by integer based algorithms in {\tt CUBE}. The PDF and CDF of $|\Delta{\bs\Psi}|$ from S256 are shown in the upper and lower axes, as functions of fine grids. Black, red, blue and green correspond to x2v2, x1v2, x2v1 and x1v1 respectively. The grey lines, marked with ``initial'', shows the distribution of actual displacement of particles in {\tt CUBEP3M}, $|{\bs\Psi}_0|$, which is orders of magnitudes larger than $|\Delta{\bs\Psi}|$. }
\label{fig.dsp}
\end{figure}

@@

\begin{figure*}
\centering
  \includegraphics[width=1.0\linewidth]{diff_power.pdf}
 \caption{Fractional deviations in the power spectrum $\Delta^2(k)/\Delta^2_0(k)-1$ (upper axes) and decorrelation $\xi$ (lower axes) induced by the integer-based algorithms in {\tt CUBE}. The four panels show the results of two redshifts in S2048S and S2048L: $\{ z=1, z=0 \} \times \{ L=400{\rm Mpc}/h, L=1200{\rm Mpc}/h \}$. The four solid lines show the results from \{x2v2,x1v2,x2v1,x1v1\}, whose colors are shown in the legends in the last panel. The orange dashed lines show the scale $k=0.2\ k_{\rm Nyquist}$, and smaller scales are beyond our interest.}
\label{fig.ccc}
\end{figure*}
\subsection{Power spectrum}
In S2048S and S2048L are two simulations with $2048^3$ particles on small ($L=400\ {\rm Mpc}/h$) and large ($L=1200\ {\rm Mpc}/h$) box sizes. We compare their accuracy by their power spectra and their cross correlations with {\tt CUBEP3M} at two checkpoints $z=1$ and $z=0$. The physical scale of S2048S is designed such that the particle mass resolution $m_p$ is comparable to TianNu and TianZero simulations (their parameters are also listed in Table \ref{t.sim}) to study halo properties. On the other hand, S2048L focuses on larger structures, on which scale one can study weak gravitational lensing, baryonic acoustic oscillation (BAO) \citep{2005ApJ...633..560E} and its reconstruction \citep{2007ApJ...664..675E,2017ApJ...841L..29W} etc.

>>>>

\shortauthors{Yu et al.} \shorttitle{CUBE}
\begin{document}
\title{CUBE: A Memory-optimized Cosmological $N$-body Algorithm}
\author{
Hao-Ran Yu$^{1,2,3}$,
Ue-Li Pen$^{2,4,5,6}$,
Xin Wang$^{2}$
}

@@

When $N$ is large, the brute force pairwise particle-particle (PP) force brings unaffordable $o(N^2)$ computations. However the workload can be alleviated to $o(N\log N)$ by many algorithms. For example, the ``tree'' algorithms like {\tt GADGET} \citep{2001NewA....6...79S,2005MNRAS.364.1105S} and its simulation ``Millennium'' \citep{2005Natur.435..629S,2012MNRAS.426.2046A}, {\tt TPM} \citep{1995ApJS...98..355X} and {\tt GOTPM} \citep{2004NewA....9..111D}, adaptive grid algorithm including {\tt HYDRA} \citep{1995ApJ...452..797C} and {\tt RAMSES} \citep{2010ascl.soft11007T}, as well as mesh-refined codes \citep{1991ApJ...368L..23C} and moving adaptive particle-mesh codes \citep{1995ApJS..100..269P}. Among these, the standard particle-mesh (PM) algorithm \citep{1988csup.book.....H} is most memory and computational efficient if we focus on large cosmological scales. The load balancing problem is minimized as the matter distribution is rather homogeneous, and the speed benefits from the fast Fourier transform (FFT) libraries, such as {\tt FFTW3} \citep{Frigo05thedesign}. {\tt PMFAST} \citep{2005NewA...10..393M} is first codes to use a 2-level PM algorithm, aiming to put PM codes toward speed, memory compactness and scalability. After subsequent developments on {\tt PMFAST}, {\tt CUBEP3M} \citep{2013MNRAS.436..540H} uses cubic spatial decomposition, and adds PP force and many other features.

Beside the methodology, the fast development of parallel supercomputers enable us to simulate a system of more than a trillion ($10^{12}$) $N$-body particles. To date the largest $N$-body simulation in application is the ``TianNu'' \citep{2017NatAs...1E.143Y,2017RAA....17...85E} run on the TianHe-2 supercomputer. With the code {\tt CUBEP3M} adding neutrino modules, it uses $3\times 10^{12}$ particles to simulate the cold dark matter (CDM) and cosmic neutrino evolution through the cosmic age.

Relative to the optimized computation optimizations, $N$-body simulations use considerable amount of memory, because the phase space coordinates $(x,y,z,v_x,v_y,v_z)$ of each $N$-body particle must be stored as at least six single precision floating point numbers (24 bytes). On the other hand, modern supercomputer systems use multi cores, many integrated cores (MIC) and even densely parallelized GPUs, bringing orders of magnitude higher computing power, whereas these architectures usually have limited memory allocation. Thus, the computation-light but memory-heavy applications, compared to matrix multiplication and decomposition calculations, are less suitable for fully usage of the computing power of modern supercomputers. For example, although native and offload modes of {\tt CUBEP3M} are able to run on the Intel Xeon-PHI MIC architectures, with the require ment of enough memory, TianNu simulation were done on TianHe-2 with only its CPUs -- 73\% of the total memory but only 13\% of the total computing power.

We present a new $N$-body simulation code {\tt CUBE}, using as low as 6 byte per particle (here after we use ``bpp'' referring to ``byte per particle''). We show that it gives accurate results in cosmological LSS simulations. In \S\ref{s.method} we show how the memory can be saved by using a ``integer-based storage'' and how the PM $N$-body algorithm is adapted with this storage format. In \S\ref{s.results} we quantify the accuracy of this algorithm by groups of simulations by {\tt CUBEP3M} and {\tt CUBE}. Discussions and conclusions are in \S\ref{s.discussion}.

\section{Method}\label{s.method}
The most memory consuming part of a $N$-body simulation is usually the phase space coordinates of $N$-body particles -- 24 bpp (4 single precision floating numbers) must be used to store each particles' 3D position and velocity vectors. {\tt CUBEP3M}, an example of a memory-efficient parallel $N$-body code, can use as low as 40 bpp in sacrificing computing speed \citep{2013MNRAS.436..540H}. This includes the phase coordinates (24 bpp) for particles in physical domain and buffered region, a linked list (4 bpp), and a global coarse mesh and local fine mesh. 4-byte real numbers are not necessarily adequate in representing the {\it global} coordinates in simulations.If the box size is many orders of magnitude larger than interactive distance between particles, especially in the field of re-simulation of dense subregions, double precision (8-byte) coordinates are needed to avoid truncation errors. Another solution is to record {\it relative} coordinates for both position and velocity. {\tt CUBE} replaces the coordinates and linked list 24+4=28 bpp memory usage with an integer based storage, reduces the basic memory usage from 28 bpp down to 6 bpp, described as following \ref{ss.position} and \ref{ss.velocity}. The algorithm is described in \ref{ss.overview}.

\subsection{Particle position storage}\label{ss.position}
We construct a uniform mesh throughout the space and each particle belongs to its parent cell of the mesh. Instead of storing global coordinates of each particle, we store its offset relative to its parent cell which contains the particle. We divide the cell, in each dimension $d$, evenly into $2^8=256$ bins, and use a 1-byte (8 bits) integer $\chi_d \in \{-128,-127,...,127\}$ to indicate which bin it locates in this dimension. The global locations of particles are given by cell-ordered format in memory space, and a complimentary number count of particle number in this mesh (density field) will give complete information of particle distribution in the mesh. Then the global coordinate in $d$th dimension $x_d$ is given by $x_d=(n_c-1)+(\chi_d+128+1/2)/256$, where $n_c=1,2,...,N_c$ is the index of the coarse grid. The mesh is chosen to be coarse enough such that the density field takes negligible memory. This coarse density field can be further compressed into 1-byte integer format, such that a 1-byte integer show the particle number in this coarse cell in range 0 to 255. In the densest cells (rarely happened) where there are $\ge 255$ particles, we can just write 255, and write the actual number as a 4-byte integer in another file.

In a simulation with volume $L^3$ and $N_c^3$ coarse cells, particle positions are stored with a precision of $L/(256N_c)$. The force calculation (e.g. softening length) should be configured much finer than this resolution, discussed in later sections. On the other hand, particle position can also be stored as 2-byte (16 bits) integers to increase the resolution. In this case, each coarse cell is divided into $2^{16}=65536$ bins and the position precision is $L/(65536N_c)$, precise enough compared to using4-byte global coordinates, see later results. We denote this case ``x2'' and denote using 1-byte integers for positions ``x1''.

@@

\begin{figure}
\centering
  \includegraphics[width=0.88\linewidth]{vdisp.pdf}
 \caption{Variance of the velocity distribution given by Equation \ref{eq.vdisp}. The $x$-axis is the scale factor characterizing the cosmic evolution and the $y$-axis shows the $\sigma_v$ in unit of km/s. The 5 curves, from top to bottom, shows the $\sigma_v(a)$ with increasing smoothing scale $r$ from 0 to 20 Mpc$/h$.}
\label{fig.vdisp}
\end{figure}

@@

The reason why we use uneven bins is that, slower particles are more abundant compared to faster ones, and one should better resolve slower particles tracing at least linear evolution. On the other hand, there could be extreme scattering particles (in case of particle-particle force), and we can safely ignore or less resolve those nonphysical particles. One of the solution is that, if we know the probability distribution function (PDF) $f(\Delta v)$ we divide its cumulative distribution function (CDF) $F(\Delta v)\in(0,1)$ into $2^{8n_\nu}$ bins to determine the boundary of $\Delta v$ bins, and particles should evenly distribute in the corresponding uneven $\Delta v$ bins. Practically we find that either $f(v_d)$ or $f(\Delta v)$ is close to Gaussian, so we can use Gaussian CDF, or any convenient analytic functions which are close to Gaussian, to convert velocity between real numbers and integers.

The essential parameter of the velocity distribution is its variance. On non-linear scale, the velocity distribution function is non-Gaussian. However, to the first order approximation, we simply assume it as Gaussian and characterized by the variance
\begin{equation}\label{eq.vdisp}
	\sigma_v(a,k) = \frac{1}{3}  (a H f D)^2 \int_0^kd^3q\frac{P_{L}(q)}{q^2},
\end{equation}
where $a(z)$ is the scale factor, $H(z)$ the Hubble parameter, $D$ is the linear growth factor, $f=d \ln D/d\ln a$, and $P_L(k)$ is the linear power spectrum of density contrast at redshift zero. $\sigma_v(a,k)$ is a function of cosmic evolution $a$ and a smoothing scale $k$, or $r$ (see Figure \ref{fig.vdisp}). $\Delta v$is the velocity dispersion relative to the coarse grid, so we approximate its variance as
\begin{equation}\label{eq.vdelta}
	\sigma^2_{\Delta}(a)=\sigma^2_v(a,r_c)-\sigma^2_v(a,r_p), 
\end{equation}
where $r_c$ is the scale of coarse grid, and $r_p$ is the scale of average particle separation. In each dimension of 3D velocity field, we use $\sigma^2_{\Delta}(a)/3$ according to the equipartition therorem. On different scales, we measure the statistics of $v_d$, $v_c$ and $\Delta v$ and find good agreement with the above model.

\begin{figure}
\centering
  \includegraphics[width=0.90\linewidth]{fig_tile.pdf}
 \caption{Spacial decomposition in {\tt CUBE} in a 2D analogy. In this example, there are 2 images per dimension (${\tt nn}=2$), and 2 tiles per image per dimension (${\tt nnt}=2$). The orange boxes show the overlapped physical+buffer=extended tile regions, inside of which one physical regions is indicated with green.}
\label{fig.tile}
\end{figure}

@@

{\tt update\_vp} -- 
The particle-mesh (PM) or particle-particle particle-mesh (P$^3$M) algorithm is applied in this subroutine to update particles's velocities.
After {\tt update\_xp}, or drift $D$, we call {\tt buffer\_density} and {\tt buffer\_xp} in order that particle positions are in buffered state, based on which, we update velocities (kick $K$) of particles in physical region of each tile. {\tt CUBE} uses a 2-level particle mesh scheme \citep{2005NewA...10..393M,2013MNRAS.436..540H}. Local fine forces have a force cutoff ${\tt nforce\_cutoff}\leq{\tt ncb}$, i.e., $F_{\rm fine}(r>{\tt ncb})=0$. So, if we apply a fine-grid particle-mesh on an extended tile with buffered-depth {\tt ncb} (see Figure \ref{fig.tile} and regard it periodic), the force on physical regions does not depend on the false periodic boundary assumption, and the resulting fine force $F_{\rm fine}$ and velocity update on physical region is correct.

@@

@@

\begin{figure*}[]
\centering
  \includegraphics[width=0.75\linewidth]{overplotparticles.pdf}
 \caption{Offset in the particle distribution induced by integer-1 based algorithm (x1v1). In S512's subregion of volume $15.625^2\times 3.90625\ ({\rm Mpc}/h)^3$, particles from {\tt CUBEP3M} (bigger grey dots in the background) and {\tt CUBE}-x1v1 (smaller red dots) are projected onto the plane of $(15.625\ {\rm Mpc}/h)^2$. The comparing rules show 1 ${\rm Mpc}/h$, fine grid and coarse grid respectively, and position resolution of x1v1 is 1/64 of a fine grid.}
\label{fig.particles}
\end{figure*}

\subsubsection{Compare with traditional algorithms}
To illustrate the improvement of memory usage, we refer to the TianNu simulation \citep{2017NatAs...1E.143Y} on the Tianhe-2 supercomputer, which used a traditional $N$-body code {\tt CUBEP3M}. TianNu's particles number is limited by memory per computing node -- for each computing node, an average of $576^3$ neutrino particles and $288^3$ CDM particles are used, and consumes about 40 GB of memory\footnote{Additional memory is used for OpenMP parallelization and for particle IDs to differentiate different particle spices.}, or about 186 bpp. A memory efficient case of {\tt CUBEP3M} by using large physical scales and at costs of speed, still uses about 40 bpp.

If the same amount of memory is allocated to {\tt CUBE}, we can set parameters as $n_\chi=n_\nu=1$, ${\tt nc}=384$, ${\tt nnt}=3$, ${\tt ncb}=6$, $\epsilon_{\rm image}=5\%$, $\epsilon_{\rm tile}=200\%$. This gives $\left\langle {\tt nplocal} \right\rangle=1536^3$, and uses about 35.6 GB of memory, corresponding to 9.84 bpp. This can be done on most of the supercomputers, even modern laptops.

@@

\subsection{Displacement of particles}
In S512, we zoom into a small region of $15.625^2\times 3.90625\ ({\rm Mpc}/h)^3$ and compare the particle distribution between {\tt CUBEP3M} and {\tt CUBE}-x1v1 in Figure \ref{fig.particles}. For clarity, {\tt CUBEP3M} particles are marked with bigger, grey dots, whereas smaller red dots are {\tt CUBE}-x1v1 particles over-plotted onto them. One can see the difference between them. The 1 ${\rm Mpc}/h$, fine grid and coarse grid scales are shown in the figure. The position resolution of particles in {\tt CUBE}-x1v1 is 1/256 of a coarse grid, or 1/64 of a fine grid.

\begin{figure}[]
\centering
  \includegraphics[width=0.99\linewidth]{ddsp_pcdf_256_box80.pdf}
 \caption{Statistics of the error in particle displacement $\Delta{\bs\Psi}$ induced by integer based algorithms in {\tt CUBE}. The PDF and CDF of $|\Delta{\bs\Psi}|$ from S256 are shown in the upper and lower axes, as functions of fine grids. Black, red, blue and green correspond to x2v2, x1v2, x2v1 and x1v1 respectively. The grey lines, marked with ``initial'', shows the distribution of actual displacement of particles in {\tt CUBEP3M}, $|{\bs\Psi}_0|$, which is orders of magnitudes larger than $|\Delta{\bs\Psi}|$. }
\label{fig.dsp}
\end{figure}

\begin{figure*}
\centering
  \includegraphics[width=1.0\linewidth]{diff_power.pdf}
 \caption{Fractional deviations in the power spectrum $\Delta^2(k)/\Delta^2_0(k)-1$ (upper axes) and decorrelation $\xi$ (lower axes) induced by the integer-based algorithms in {\tt CUBE}. The four panels show the results of two redshifts in S2048S and S2048L: $\{ z=1, z=0 \} \times \{ L=400{\rm Mpc}/h, L=1200{\rm Mpc}/h \}$. The four solid lines show the results from \{x2v2,x1v2,x2v1,x1v1\}, whose colors are shown in the legends in the last panel. The orange dashed lines show the scale $k=0.2\ k_{\rm Nyquist}$, and smaller scales are beyond our interest.}
\label{fig.ccc}
\end{figure*}
\subsection{Power spectrum}

@@

In S2048S and S2048L are two simulations with $2048^3$ particles on small ($L=400\ {\rm Mpc}/h$) and large ($L=1200\ {\rm Mpc}/h$) box sizes. We compare their accuracy by their power spectra and their cross correlations with {\tt CUBEP3M} at two checkpoints $z=1$ and $z=0$. The physical scale of S2048S is designed such that the particle mass resolution $m_p$ is comparable to TianNu and TianZero simulations (their parameters are also listed in Table \ref{t.sim}) to study halo properties. On the other hand, S2048L focuses on larger structures, on which scale one can study weak gravitational lensing, baryonic acoustic oscillation (BAO) \citep{2005ApJ...633..560E} and its reconstruction \citep{2007ApJ...664..675E,2017ApJ...841L..29W} etc.
